FraCaS: 80 percent complete

### 2019-02-26 paper2018/paper.tex
compress obviously compressible things

\begin{abstract}
  TODO
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches. Consider the following
two examples, one from the SNLI dataset, predominatly used to test DL
approaches and which has a state-of-the-art performace of 0.911
\citet{liu:2019}, and the FraCaS test suite, the dataset \textit{par
  excellence} to test logical approaches.
%
The following two examples illustrate the issue.
\begin{lingex}
\item % An SNLI example \\
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
\item % An example from FraCaS\\
\textbf{P1}] A Scandinavian won the Nobel Prize.\\
\textbf{P2}	Every Swede is  Scandinavian.\\
\textbf{Q}  Did a Swede win the Nobel prize? \\
\textbf{H} A Swede won the Nobel prize.\\
\textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In the SNLI example, one has to make a number of non-trivial
assumptions in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
and, thus, decide whether one description contradicts the other, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning can be also performed. Such an example is the FraCaS
example, as well as very fine-grained cases like the Antecedent
Contained Deletion ellipsis example below:

\begin{lingex}

\item An example from FraCaS involving fine-grained reasoning\\
\textbf{P1}	Bill spoke to everyone that John did.	\\	
\textbf{P2}	John spoke to Mary.	\\
\textbf{Q} 	Did Bill speak to Mary?\\
\textbf{H} 	Bill spoke to Mary.\\
\textbf{Label:	UNK} [FraCaS 065] 	\end{lingex}

@@

In this paper, we concentrate on this sort of fine-grained, logical
reasoning. In particular, we present the first system that deals with
the sections on ellipsis and anaphora in the FraCaS test suite and has
both the best coverage, as well as accuracy compared to other logical
approaches. More specifically, the contributions of the paper are:

\begin{itemize}
\item A two-phase system (rather than a one-phase as described by
  \citet{bernardy_type_2017}): a) using Haskell as a converter between
  Grammatical Framework trees and Coq types, b) using various type
  combinators to reach the desired NL semantics.
\item The first run of the Anaphora and Ellipsis sections of the
  FraCaS (based on work by \citet{bernardy_jolli})
 \item  More coverage than any previous logical approach
 \item State-of-the-art results in terms of accuracy compared to all
   previous approaches, for every section of the FraCaS
  \item Handing of multiple readings
  \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
  \item New handling of prepositions and adverbs compared to the
    earlier system in \citet{bernardy_type_2017}, as well as new
    handling of definitions, enabled by the dynamic semantics system.
  \item Improvement over the previous treatment of Plurals/Quantifiers
  \item Big improvement on the comparatives section enabled by
    handling comparatives using a threshold (BUT one different to the
    one used for adjectives). Again, the dynamic semantics system is
    crucial here.

\end{itemize}
\section{Background: GF, Monadic Dynamic Semantics, Coq}

\section{GF}

@@

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program, which implements a dynamic
  semantics and comprises the bulk of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. Such aspects include for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}
The architecture is represented schematically in Figure \ref{fig:overview}.

\begin{figure*}
  \centering
{
  \tiny
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
                    semithick]
  \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]

  \node[state]         (txt)                {text};
  \node[state]         (ast) [right of=txt] {AST};
  \node[state]         (pp)  [right of=ast] {Probabilistic program};
  \node[state]         (res) [right of=pp] {result in $(0,1)$};

  \path (txt) edge              node (GF)  {GF}                  (ast)
        (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
        (pp) edge               node (Coq) {Coq} (res);

  \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
  \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
  \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};

  \path
  (Syntax) edge (GF)
  (Sem) edge (Hask)
  (Hyper) edge (Coq);
\end{tikzpicture}
}
\caption{Phases in our system. ($\ast$) At the level of Coq, we handle
  the details of the adverbial (veridicality properties) and
  adjectival semantics (division into subsective, extentional,
  non-committal, etc. categories.) }
  \label{fig:overview}
\end{figure*}

@@

\paragraph{anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples such as the following:
\begin{lingex}
\item[FraCaS 122]
P1	Every committee has a chairman.\\
P2	He is appointed its members.\\
H 	Every committee has a chairman appointed by members of the committee. 
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase. Such
ellptic expressions are handled in the same way as anaphora.
\begin{lingex}
\item[FraCas 173]
  P1	Bill spoke to everyone that John did [elliptic V2].\\
  P2	John spoke to Mary.\\
  Q 	Did Bill speak to Mary?\\
  H 	Bill spoke to Mary.
\end{lingex}
For example, in (FraCaS 173, P1), the argument of ``did'' is
explicitly marked as an elliptic V2, which we resolve to ``speak'' in
that context.
\paragraph{definites} A naive way to handle definites is an
existential (or sigma). However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

@@

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions. Finally, ``more'' is intepreted as ``at least
<elliptic quantity>+1''. This treatment, which we illustrated here on
an example, is systematic in our implementation.

@@

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by Bill MacCartney\todo{cite} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labelling (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of the cases listed in Table \ref{tab:overrulings}, which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.
\begin{table*}
  \centering
\begin{tabular}{ccc}
 test case & new class & comment \\ \hline
    005 & UNK & missing hypothesis: there are italian tenors \\
   056 &  Yes & Already identified as such by MacCartney \\
  069 & Unk & Mary could have used someone else's workstation \\
  119 & Unk & \textit{ibid.} \\
  181 & Yes & for the same reason as 180 \\
  226 & Yes &
\end{tabular}
  \caption{Overruled FraCaS cases}
  \label{tab:overrulings}
\end{table*}

>>>>

\begin{abstract}
  TODO
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches. Consider the following
two examples, one from the SNLI dataset, predominatly used to test DL
approaches and which has a state-of-the-art performace of 0.911
\citet{liu:2019}, and the FraCaS test suite, the dataset \textit{par
  excellence} to test logical approaches.
%
\begin{lingex}
\item % An SNLI example \\
  \label{ex:snli}
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
% \item % An example from FraCaS\\
% \textbf{P1} A Scandinavian won the Nobel Prize.\\
% \textbf{P2}	Every Swede is  Scandinavian.\\
% \textbf{Q}  Did a Swede win the Nobel prize? \\
% \textbf{H} A Swede won the Nobel prize.\\
% \textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In \ref{ex:snli}, one has to make a number of non-trivial
assumptions in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
and, thus, decide whether one description contradicts the other, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning can be also performed. Such an example is the FraCaS
example, as well as very fine-grained cases like the Antecedent
Contained Deletion ellipsis example \ref{ex:acd} below.\todo{It'd be nice to have a more clear parallel with indefinite articles as well; but I could not find this in fracas.}

% \begin{lingex}
% \item % An example from FraCaS involving fine-grained reasoning\\
% \textbf{P1}	Bill spoke to everyone that John did.	\\	
% \textbf{P2}	John spoke to Mary.	\\
% % \textbf{Q} 	Did Bill speak to Mary?\\
% \textbf{H} 	Bill spoke to Mary.\\
% \textbf{Label:	UNK} [FraCaS 065] 	\end{lingex}

@@

In this paper, we concentrate on this sort of fine-grained, logical
reasoning. In particular, we present a logic-based system that deals
with many linguistic phenomena \emph{at the same time}. It is the
first systems covering the sections on ellipsis and anaphora in the
FraCaS test suite and has the best coverage and accuracy on the
overall test suite.
% This is listed below; we can let the reader figure it out.
% \begin{itemize}
% \item A two-phase system (rather than a one-phase as described by
%   \citet{bernardy_type_2017}): a) using Haskell as a converter between
%   Grammatical Framework trees and Coq types, b) using various type
%   combinators to reach the desired NL semantics.
% \item The first run of the Anaphora and Ellipsis sections of the
%   FraCaS (based on work by \citet{bernardy_jolli})
%  \item  More coverage than any previous logical approach
%  \item State-of-the-art results in terms of accuracy compared to all
%    previous approaches, for every section of the FraCaS
%   \item Handing of multiple readings
%   \item  Threshold-based interpretation of (some) adjectives plus Linear arithmetic proofs
%   \item New handling of prepositions and adverbs compared to the
%     earlier system in \citet{bernardy_type_2017}, as well as new
%     handling of definitions, enabled by the dynamic semantics system.
%   \item Improvement over the previous treatment of Plurals/Quantifiers
%   \item Big improvement on the comparatives section enabled by
%     handling comparatives using a threshold (BUT one different to the
%     one used for adjectives). Again, the dynamic semantics system is
%     crucial here.
% \end{itemize}

\section{Background: GF, Monadic Dynamic Semantics, Coq}

@@

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program, which implements a dynamic
  semantics and comprises the bulk of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. Such aspects include for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}

% The architecture is represented schematically in Figure \ref{fig:overview}.
% \begin{figure*}
%   \centering
% {
%   \tiny
% \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.7cm,
%                     semithick]
%   \tikzstyle{every state}=[draw=black,text=black,shape=rectangle,text width=2cm]

%   \node[state]         (txt)                {text};
%   \node[state]         (ast) [right of=txt] {AST};
%   \node[state]         (pp)  [right of=ast] {Probabilistic program};
%   \node[state]         (res) [right of=pp] {result in $(0,1)$};

%   \path (txt) edge              node (GF)  {GF}                  (ast)
%         (ast) edge              node (Hask) [text width=1.5cm]{Haskell interpreter} (pp)
%         (pp) edge               node (Coq) {Coq} (res);

%   \node (Syntax) [above of=GF,node distance=1cm] {Syntax};
%   \node (Sem) [above of=Hask,node distance=1cm] {Dynamic Semantics};
%   \node (Hyper) [above of=Coq,node distance=1cm] {Additional Semantics$^\ast$};

%   \path
%   (Syntax) edge (GF)
%   (Sem) edge (Hask)
%   (Hyper) edge (Coq);
% \end{tikzpicture}
% }
% \caption{Phases in our system. ($\ast$) At the level of Coq, we handle
%   the details of the adverbial (veridicality properties) and
%   adjectival semantics (division into subsective, extentional,
%   non-committal, etc. categories.) }
%   \label{fig:overview}
% \end{figure*}

@@

\paragraph{anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples such as the following:
\begin{lingex}
\item
    \label{ex:anaphora-scope}
\textbf{P1}	Every committee has a chairman.\\
\textbf{P2}	He is appointed its members.\\
\textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase. Such
ellptic expressions are handled in the same way as anaphora.
\begin{lingex}
\item
    \label{ex:acd}
\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
\textbf{P2}	John spoke to Mary.\\
\textbf{Q} 	Did Bill speak to Mary?\\
\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
\end{lingex}
For example, in (FraCaS 173, P1), the argument of ``did'' is
explicitly marked as an elliptic V2, which we resolve to ``speak'' in
that context.
\paragraph{definites} A naive way to handle definites is an
existential (or sigma). However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

@@

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ $\ge$ 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions. Finally, ``more'' is intepreted as ``at least
<elliptic quantity>+1''. This treatment, which we illustrated here on
an example, is systematic in our implementation.

@@

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by Bill MacCartney\todo{cite} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labelling (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of the cases listed in Table \ref{tab:overrulings}, which appeared to have been
misclassified. We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.
\begin{table}
  \centering
  \small
\begin{tabularx}{\columnwidth}{ccp{3.8cm}}
 test case & new class & comment \\ \hline
    005 & UNK & missing hypothesis: there are italian tenors \\
   056 &  Yes & Already identified as such by MacCartney \\
  069 & Unk & Mary could have used someone else's workstation \\
  119 & Unk & \textit{ibid.} \\
  181 & Yes & for the same reason as 180 \\
  226 & Yes &
\end{tabularx}
  \caption{Overruled FraCaS cases}
  \label{tab:overrulings}
\end{table}
